Spin canonical rings of log stacky curves

### 2015-07-08 curves-write-up/spin_cring.tex
edits for more meaty comment, mid edit

\todo{Think about John's comment to regrade the ring.}
\todo{Brian: Don't write out author's name every time.}

@@

In order to find generators and relations for $M(\Gamma)$, we
translate the seemingly analytic question of understanding the ring
of modular forms into the algebraic category, using a
generalization of the GAGA principle.
\todo{brain: The appeal to stacky GAGA on page 1 sweeps a bit under the rug.  I suppose you noticed (as I just did upon glancing at [14] on the arxiv) that the proof of Prop. 6.1.5 in [14] seems to pass in silence over the removal of the properness hypothesis (which, in the scheme case — no stacks — requires using something like the little Picard theorem or some other device to deal with avoidance of essential singularities).  That is moot for you since only the proper case is relevant for your motivation.  More importantly, since the entire discussion is about spaces of sections of coherent sheaves, one should address stacky GAGA at the level of such $H^0$’s, which is not part of Prop. 6.1.5 in [14], to have adequately treated the motivation.  On line -1 of last paragraph on page 1, should say X is proper and geometrically connected over k.  Note that the proof of Lemma 10.2.1 in [14] gives a thoroughly analytic construction of the line bundle (no ad hoc stuff with divisors), so to link it back to the algebraic case one is using an honest GAGA comparison at the level of categories of coherent sheaves too (with refinement for line bundles).}
 By \cite[Proposition 6.1.5]
{vzb:stacky}, there is an equivalence of categories between orbifold
curves and log stacky curves over $\BC$. Hence, for the remainder
of the paper, we will work in the algebraic category.

@@

If there is some odd $k$ for which $M_k(\Gamma) \neq 0$, then the
ring of modular forms $M(\Gamma)$ is generated as a $\BC$-algebra
by elements of degree at most $\max(5, e_1, \ldots, e_r)$ with
relations generated in degree at most $2 \cdot \max(5, e_1, \ldots, 
e_r)$ so long as $\sigma$ does not lie in a finite list of
exceptional cases which are listed and described in Table
~\ref{table:g-1-exceptional} for signatures with $g = 1$ and in
Table ~\ref{table:g-0-exceptional} for signatures with $g = 0$.
\end{cor}
\todo{Add somewhere that if there is an odd $k$ with nonzero weight, it will generally have to occur in weight 3, according to brian 2.}
\begin{rem}
\label{rem:factor-two}
In the case that $M_k(\Gamma) = 0$ for all odd $k$, the generation bound of $6 \cdot \max(3, e_1, \ldots, e_r)$ and relation
bound of $12 \cdot \max(3, e_1, \ldots, e_r)$ can be reduced to $2\cdot \max(3,e_1, \ldots, e_r)$ and $4 \cdot \max(3,e_1, \ldots, e_r)$, apart from several small families of cases. See 
~\cite[Theorem 9.3.1]{vzb:stacky} and ~\cite[Theorem 8.7.1]{vzb:stacky} for a more precise statement of these bounds in the cases that $g = 0$ and $g > 0$ respectively. Note that we multiply all bounds given in Voight and Zureick-Brown \cite{vzb:stacky} by a factor of two, because our grading convention on $\bigoplus_k M_k(\Gamma)$ differs from that of Voight and Zureick-Brown by a factor of 2.
\end{rem}

\todo{Add example saying all congruence subgroups with a form of odd weight are generated in weight at most 5 with relations in weight at most 12, (which can be reduced to 10 if the signature is not (0;3,3,3,3;0). Further, if $L_X$ is effective, then generation will be in weight at most 3.}

@@

\begin{defn}
\label{defn:stacky-curve}
A \textbf{stacky curve} $\sx$ over an algebraically closed field $k$ is the datum of a smooth proper connected scheme $X$ of dimension $1$, together with a finite number of closed points of $X$, $P_1, \ldots, P_r$, called {\bf stacky points}, with {\bf stabilizer orders} $e_1, \ldots, e_r \in \BZ_{\geq 2}.$ The scheme $X$ associated to a stacky curve $\sx$ is called the {\bf coarse space} of $\sx$.

@@

In this section, we present several directions for further research.

\begin{enumerate}
	\item As noted in Remark ~\ref{rem:explicit-generators}, the 
		proof of Theorem ~\ref{thm:main} 
		%given in this paper 
		gives an 
		explicit procedure for computing the generators and relations of $R_
		\halfcan$ when the genus of $X$ is $0$ or $1$.
		\begin{question}
		\label{ques:algorithm}
			Is there an explicit procedure for computing the generators 
			and relations for an arbitrary genus $g \geq 2$ curve?
		\end{question}
		Suppose the generators and relations of $R_{\halfcan_X}$ are known, 
		where $(\sx, \Delta, L)$ is a spin canonical ring, $X$ is the 
		coarse space of $\sx$, and $L_X \leq L$ is a spin canonical divisor 
		of $X$. In this case our proof of Theorem ~\ref{thm:main} yields an 
		algorithm for inductively determining the generators of $\halfcan$. 
		This reduces Question \ref{ques:algorithm} to the 
		following Petri-like question:
		\begin{question}
		\label{ques:spin-canonical-petri}
			Given a {\rm(}non-stacky {\rm)} curve $X$ and a spin canonical 
			divisor $L \in \di X$, can one write down a set of minimal 
			generators and minimal relations for $R_\halfcan$?
		\end{question}

@@

\section{Acknowledgments}
We are grateful to David Zureick-Brown for introducing us to the
study of stacky canonical rings, for providing invaluable guidance,
and for his mentorship. We also thank Ken Ono and the
Emory University Number Theory REU for arranging our project and
providing a great environment for mathematical learning and
collaboration. We thank Jorge Sentieiro Neves for pointing out the
bounds given by Miles Reid's work on the maximal degrees of minimal
generators and relations of spin canonical rings in high genera.
We would also like to mention Evan O'Dorney for enlightening
conversations regarding his work in this field.
Finally, we gratefully acknowledge the financial support given by
NSF Grant Award Number 1250467 via the Emory University Number
Theory REU. We deeply appreciate all of the support that has made
our work possible.

>>>>

\todo{Think about John's comment to regrade the ring.}

@@

In order to find generators and relations for $M(\Gamma)$, we
translate the seemingly analytic question of understanding the ring
of modular forms into the algebraic category, using a
generalization of the GAGA principle.
 By \cite[Proposition 6.1.5]
{vzb:stacky}, there is an equivalence of categories between orbifold
curves and log stacky curves over $\BC$. Hence, for the remainder
of the paper, we will work in the algebraic category.

@@

If there is some odd $k$ for which $M_k(\Gamma) \neq 0$, then the
ring of modular forms $M(\Gamma)$ is generated as a $\BC$-algebra
by elements of degree at most $\max(5, e_1, \ldots, e_r)$ with
relations generated in degree at most $2 \cdot \max(5, e_1, \ldots, 
e_r)$ so long as $\sigma$ does not lie in a finite list of
exceptional cases which are listed and described in Table
~\ref{table:g-1-exceptional} for signatures with $g = 1$ and in
Table ~\ref{table:g-0-exceptional} for signatures with $g = 0$.
\end{cor}

\begin{rem}
\label{rem:gen-at-most-three}
In the case that $M(\Gamma)$ has some odd weight modular form, it in fact has an odd weight modular form in weight $3$. This holds in the case that the genus $g \geq 2$ because $\deg \lfloor 3L \rfloor > 2g-1,$ so by Riemann--Roch, $\dim_k H^0(\sx,3L) > 0,$ implying there must be some generator in weight $1$ or $3$. The case that $M(\Gamma)$ has a generator in degree 1 or 3 when $g(X) = 0$ or $1$ follows from the fact that it has such a generator in the base cases given in Tables ~\ref{table:g-0-base-cases} and ~\ref{table:g-1-base}. Hence, the bound on the degree of generators and relations in the case that $M(\Gamma)$ has some odd weight modular form, given in Corollary ~\ref{cor:main-mod-forms}, is closely related to the degree of the minimal odd weight modular form.
\end{rem}

\todo{Add example saying all congruence subgroups with a form of odd weight are generated in weight at most 5 with relations in weight at most 12, (which can be reduced to 10 if the signature is not (0;3,3,3,3;0). Further, if $L_X$ is effective, then generation will be in weight at most 3.}
\begin{example}
\label{eg:congruence-bounds}
Suppose $\Gamma \subset SL_2(\BZ)$ is a congruence subgroup. Then, since the action  of $SL_2(\BZ)$ has points with stabilizer order 2 and 3, the action of $\Gamma$ can only have points with stabilizer order 2 and 3. Furthermore, if $\Gamma$ has some odd weight modular form, then it must have no points with stabilizer order 2 (see Remark ~\ref{rem:odd-denom}). Therefore, by Corollary ~\ref{cor:main-mod-forms}, $M(\Gamma)$ is generated in degree at most $5$ with a relation in degree at most $12$. Furthermore, if the coarse space associated $\Gamma$ does not have .
\end{example}

\begin{rem}
\label{rem:factor-two}
In the case that $M_k(\Gamma) = 0$ for all odd $k$, the generation bound of $6 \cdot \max(3, e_1, \ldots, e_r)$ and relation
bound of $12 \cdot \max(3, e_1, \ldots, e_r)$ can be reduced to $2\cdot \max(3,e_1, \ldots, e_r)$ and $4 \cdot \max(3,e_1, \ldots, e_r)$, apart from several small families of cases. See 
~\cite[Theorem 9.3.1]{vzb:stacky} and ~\cite[Theorem 8.7.1]{vzb:stacky} for a more precise statement of these bounds in the cases that $g = 0$ and $g > 0$ respectively. Note that we multiply all bounds given in Voight and Zureick-Brown \cite{vzb:stacky} by a factor of two, because our grading convention on $\bigoplus_k M_k(\Gamma)$ differs from that of Voight and Zureick-Brown by a factor of 2.
\end{rem}

@@

\begin{defn}
\label{defn:stacky-curve}
A \textbf{stacky curve} $\sx$ over an algebraically closed field $k$ is the datum of a smooth proper integral scheme $X$ of dimension $1$, together with a finite number of closed points of $X$, $P_1, \ldots, P_r$, called {\bf stacky points}, with {\bf stabilizer orders} $e_1, \ldots, e_r \in \BZ_{\geq 2}.$ The scheme $X$ associated to a stacky curve $\sx$ is called the {\bf coarse space} of $\sx$.

@@

In this section, we present several directions for further research.
\todo{Edit to eliminate John's confusions}
\begin{enumerate}
	\item As noted in Remark ~\ref{rem:explicit-generators}, the 
		proof of Theorem ~\ref{thm:main} 
		%given in this paper 
		gives an 
		explicit procedure for computing the generators and relations of $R_
		\halfcan$ when the genus of $X$ is $0$ or $1$.
		\begin{question}
		\label{ques:algorithm}
			Is there an explicit procedure for computing the generators 
			and relations for an arbitrary genus $g \geq 2$ curve?
		\end{question}
		Suppose the generators and relations of $R_{\halfcan_X}$ are known, 
		where $(\sx, \Delta, L)$ is a spin canonical ring, $X$ is the 
		coarse space of $\sx$, and $L_X \leq L$ is a spin canonical divisor 
		of $X$. In this case our proof of Theorem ~\ref{thm:main} yields an 
		algorithm for inductively determining the generators of $\halfcan$. 
		This reduces Question \ref{ques:algorithm} to the 
		following Petri-like question:
		\begin{question}
		\label{ques:spin-canonical-petri}
			Given a {\rm(}non-stacky {\rm)} curve $X$ and a spin canonical 
			divisor $L \in \di X$, can one write down a set of minimal 
			generators and minimal relations for $R_\halfcan$?
		\end{question}

@@

\section{Acknowledgments}
We are grateful to David Zureick-Brown for introducing us to the
study of stacky canonical rings, for providing invaluable guidance,
and for his mentorship. We also thank Ken Ono and the
Emory University Number Theory REU for arranging our project and
providing a great environment for mathematical learning and
collaboration. We thank Jorge Sentieiro Neves for pointing out the
bounds given by Miles Reid's work on the maximal degrees of minimal
generators and relations of spin canonical rings in high genera.
We thank Brian Conrad and John Voight for helpful comments on this paper. The first author would also like to mention Joseph Harris and Evan O'Dorney for several enlightening conversations related to Petri's Theorem.
Finally, we gratefully acknowledge the financial support given by
NSF Grant Award Number 1250467 via the Emory University Number
Theory REU. We deeply appreciate all of the support that has made
our work possible.
